---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---
```{r, echo=TRUE}

```
## Load necessary libraries
```{r, echo=TRUE}
library(dplyr)
library(lubridate)
library(RColorBrewer)
```

## prepare a color palette
```{r}
accent <- brewer.pal(8,"Accent")
```

## Loading and preprocessing the data
```{r}
data <- read.csv("activity.csv")
data <- mutate(data, date = ymd(date))
```

## What is mean total number of steps taken per day?
### first get the total number of steps per day
```{r}
by_date <- group_by(data,date)
total <- summarise(by_date,total_steps = sum(steps))
```
### plot a histogram
```{r}
hist(total$total_steps, main="Histogram Steps/Day",xlab="Steps",breaks=10,col=accent[5])
```

### calculate the mean and median
```{r}
mean(total$total_steps,na.rm=TRUE)
median(total$total_steps,na.rm=TRUE)
```





## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?


